---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Fatimah Yunusa
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false
#install.packages("arrow")
#install.packages("tidyverse")
#install.packages("haven")
#install.packages("modelsummary")
#install.packages("kableExtra")
#install.packages("rstanarm")
#install.packages("bayesplot")
#install.packages("ggplot2")
#install.packages("brant")
#install.packages("pscl")

library(arrow)
library(tidyverse)
library(haven)
library(tidyverse)
library(palmerpenguins)
library(modelsummary)
library(kableExtra)
library(rstanarm)
library(bayesplot)
library(ggplot2)
library(brant)
library(pscl)



```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as
follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data
[@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an
entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome
variable or update the subheading to be singular.

```{r}



```



## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine
a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to
investigate... Background details and diagnostics are included in
[Appendix -@sec-model-details].

## Model set-up


# Results

Our results are summarized in 

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false



```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The
discussion is a chance to show off what you know and what you learnt
from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what
your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Diagnostics



\newpage

# References
