---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Fatimah Yunusa
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false

#install.packages("rstanarm")
#install.packages("modelsummary")
library(modelsummary)
library(rstanarm)



```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as
follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data
[@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an
entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome
variable or update the subheading to be singular.




## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine
a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to
investigate... Background details and diagnostics are included in
[Appendix -@sec-model-details].

## Model set-up


# Results

Our results are summarized in @tbl-modelresults

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

bayesian_model <- readRDS(file = here::here ("models/bayesian_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Summary of the Ordinal Logistic regression model"
#| warning: false


# Install necessary packages if not already installed
#install.packages("broom.mixed")
#install.packages("kableExtra")

# Load the packages
library(broom.mixed)
library(kableExtra)

# Generate a tidy summary of your Bayesian model
tidy_summary <- tidy(
  bayesian_model, 
  conf.int = TRUE,      # Include credible intervals
  conf.level = 0.95     # Set confidence level to 95%
)

# Render the summary table as a nice HTML/LaTeX table
tidy_summary %>%
  kable(
    caption = "Bayesian Model Summary with Credible Intervals",
    digits = 2,             # Round numbers to 2 decimal places
    col.names = c("Parameter", "Estimate", "Std. Error", "Lower 95% CI", "Upper 95% CI")  # Rename columns
  ) %>%
  kable_styling(
    full_width = FALSE,
    bootstrap_options = c("striped", "hover", "condensed")
  )




```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The
discussion is a chance to show off what you know and what you learnt
from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what
your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Diagnostics



\newpage

# References
